```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Package "eqplot"

This package will provide tools to visualize the earthquake dataset obtained from the U.S. National Oceanographic and Atmospheric Administration (NOAA) on significant earthquakes around the world. This dataset contains information about 5,933 earthquakes over an approximately 4,000 year time span.


## Installation

You can install eqplot from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("atibot/eqplot", build_vignettes = TRUE)
```


## Examples

These are basic examples which show you how to solve common problems. 

### Timeline plot

```{r example_timeline, eval = FALSE}
## basic example code
data("raw_df")
test_df <-
raw_df %>% eq_clean_data() %>%
dplyr::filter(COUNTRY == "TURKEY" | COUNTRY == "JAPAN") %>%
dplyr::filter(DATE >= eq_date("2000-01-01") & DATE <= eq_date("2017-01-01"))

ggtimeline <- ggplot2::ggplot(test_df)
ggtimeline2 <- ggtimeline + geom_timeline(
ggplot2::aes(
x = DATE,
y = COUNTRY,
size = EQ_PRIMARY,
color = TOTAL_DEATHS,
fill = TOTAL_DEATHS
)
) +
eq_theme +
ggplot2::scale_size_continuous(name = "Richer scale value", breaks = c(0, 2, 4, 6, 8)) +
ggplot2::scale_color_continuous(name = "# deaths") +
ggplot2::scale_fill_continuous(name = "# deaths") +
ggplot2::labs(title = "Earthquakes")

ggtimeline3 <- ggtimeline2 + geom_timeline_label(ggplot2::aes(
x = DATE,
y = COUNTRY,
mag = EQ_PRIMARY,
label = LOCATION_NAME
),
n_max = 5)

ggtimeline3
```

### Map plot with Date 

```{r example_map1, eval = FALSE}
data("raw_df")
raw_df %>% eq_clean_data() %>%
dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>%
eq_map(annot_col = "DATE")
```

### Map plot with Location, Magnitude and Deaths

```{r example_map2, , eval = FALSE}
data("raw_df")
raw_df %>% eq_clean_data() %>%
dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>%
dplyr::mutate(popup_text = eq_create_label(.)) %>%
eq_map(annot_col = "popup_text")
```

## Vignette

Once the package is installed, view the vignette for more details.

```{r vignette, eval= FALSE}
vignette(topic = "eqplot", package = "eqplot")
```
